% Cover letter — Hp3 manuscript, Microbial Biotechnology (Wiley)
% Build:  latexmk cover_letter.tex   (xelatex)
\documentclass[11pt,a4paper]{article}
\usepackage{iftex}
\ifPDFTeX\usepackage[T1]{fontenc}\usepackage[utf8]{inputenc}
\else\usepackage{fontspec}\setmainfont{Times New Roman}\fi
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{parskip}
\usepackage{hyperref}\hypersetup{hidelinks}
\pagestyle{empty}

\begin{document}

\noindent Jinlong Ru, on behalf of all authors\\
Technical University of Munich \& Helmholtz Centre Munich\\
\href{mailto:jinlong.ru@tum.de}{jinlong.ru@tum.de}

\bigskip
\noindent The Editor-in-Chief\\
\emph{Microbial Biotechnology}

\bigskip
\noindent Dear Editor,

We are pleased to submit our manuscript, \textbf{``A deeply divergent
\emph{Aeromonas} bacteriophage rescues largemouth bass from lethal
\emph{Aeromonas hydrophila} infection,''} for consideration as a Research
Article in \emph{Microbial Biotechnology}.

\emph{Aeromonas hydrophila} is a leading aquaculture pathogen whose escalating
multidrug resistance demands sustainable, non-antibiotic control strategies.
We report the isolation and multi-level characterization of bacteriophage
vB\_AhyM\_Hp3 (Hp3) against \emph{A. hydrophila}, with three findings we
believe are of direct interest to the journal's biocontrol readership:

\begin{enumerate}
\item \textbf{A deeply divergent new viral genus.} Integrating nucleotide-level
(VIRIDIC, 83.65\% intergenomic similarity), whole-proteome (ViPTree),
amino-acid genome-BLAST distance phylogeny (VICTOR), and protein-sharing
network (vConTACT3) analyses, Hp3 and its closest relative HJ05 form a new,
deeply divergent genus within the class \emph{Caudoviricetes}---with no more
than $S_G = 0.073$ whole-proteome similarity to any classified virus and clear
separation from all established families, suggesting a putative new
family-level lineage pending broader sampling and formal ICTV evaluation.

\item \textbf{Demonstrated \emph{in vivo} efficacy and host safety.} A single
administration (MOI 1.0) increased the seven-day survival of largemouth bass
from 15\% to 85\% under lethal challenge, while fish receiving phage alone
showed 100\% survival with no histopathology.

\item \textbf{A genomically safe, well-defined biocontrol candidate.} Hp3
carries no antibiotic-resistance genes and no high-confidence virulence
determinants, and exhibits rapid lytic kinetics with broad temperature and pH
stability.

\end{enumerate}

Together these results position Hp3 both as a practical biocontrol agent for
aquaculture and as a foundation for next-generation phage-based therapeutics
and cocktail formulations --- squarely within the scope of
\emph{Microbial Biotechnology}.

The manuscript is original, has not been published previously, and is not under
consideration elsewhere. As corresponding author, I consent to its publication
and confirm that the content and authorship of this manuscript have been
approved by all authors; the authors declare no competing interests. All raw sequencing data and assembled genomes are
openly available in NCBI BioProject \textbf{PRJNA1348731}; figure source data
are deposited in Figshare
(\href{https://doi.org/10.6084/m9.figshare.30500525}{10.6084/m9.figshare.30500525})
and the analysis code is archived in Zenodo
(\href{https://doi.org/10.5281/zenodo.17454899}{10.5281/zenodo.17454899}).
All animal experiments followed the ARRIVE guidelines and were approved by the
Institutional Animal Care and Use Committee (approval no. NWAFU-314020038).

% --------------------------------------------------------------------------
% Suggested reviewers — five experts covering the three pillars of the paper
% (phage genomics/taxonomy; Aeromonas/aquaculture phage therapy; fish in-vivo
% disease models). None are collaborators or co-authors of the authors.
% Emails verified from institutional / corresponding-author pages (2026-05).
% Substitute bench: Dann Turner (UWE Bristol, Dann2.Turner@uwe.ac.uk) for a
% pillar-1 decline.
% --------------------------------------------------------------------------
\medskip
\noindent\textbf{Suggested reviewers} (none has a conflict of interest; none is a
recent collaborator or co-author of the authors):
\begin{itemize}
\item \textbf{Evelien M. Adriaenssens}, Quadram Institute Bioscience, Norwich, UK
(\href{mailto:evelien.adriaenssens@quadram.ac.uk}{evelien.adriaenssens@quadram.ac.uk})
--- bacteriophage genomics and prokaryotic virus taxonomy; Chair of the ICTV
Bacterial Viruses Subcommittee (\emph{Caudoviricetes} classification framework).
\item \textbf{Andrew D. Millard}, University of Leicester, UK
(\href{mailto:adm39@leicester.ac.uk}{adm39@leicester.ac.uk}) --- bacteriophage
comparative genomics, reference-genome curation, and therapeutic phages.
\item \textbf{Mao Lin}, Fisheries College, Jimei University, Xiamen, China
(\href{mailto:linmao@jmu.edu.cn}{linmao@jmu.edu.cn}) --- \emph{Aeromonas} and
aquaculture bacteriophage genomics and phage therapy (host-range and genome
characterization).
\item \textbf{Jianguo Su}, College of Fisheries, Huazhong Agricultural University,
Wuhan, China (\href{mailto:sujianguo@mail.hzau.edu.cn}{sujianguo@mail.hzau.edu.cn})
--- fish antibacterial immunity and \emph{A.~hydrophila} \emph{in vivo} challenge
models.
\item \textbf{Lingbing Zeng}, Yangtze River Fisheries Research Institute, Chinese
Academy of Fishery Sciences, Wuhan, China
(\href{mailto:zenglingbing@gmail.com}{zenglingbing@gmail.com}) --- fish pathology
and largemouth bass (\emph{Micropterus salmoides}) disease models.
\end{itemize}

\medskip
\noindent We thank you for considering our work and look forward to your
response.

\bigskip
\noindent Sincerely,\\[1ex]
Jinlong Ru, Li Deng, and Gaoxue Wang\\
\emph{(corresponding authors)}\\
on behalf of all authors

\end{document}
