% Cover letter — Hp3 manuscript, Cell Press MULTI-JOURNAL SUBMISSION (Cell Reports + iScience)
% Build:  latexmk -pdfxe cover_letter-mjs.tex   (xelatex)
% Supersedes cover_letter-cr.tex (single-CR) per DEC-D027. Differs in addressee
% (Multi-Journal Submission) and a paragraph framing fit for BOTH Cell Reports
% (broad interest) and iScience (rigorous, multidisciplinary, sound science).
\documentclass[11pt,a4paper]{article}
\usepackage{iftex}
\ifPDFTeX\usepackage[T1]{fontenc}\usepackage[utf8]{inputenc}
\else\usepackage{fontspec}\setmainfont{Times New Roman}\fi
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{parskip}
\usepackage{hyperref}\hypersetup{hidelinks}
\pagestyle{empty}

\begin{document}

\noindent Jinlong Ru, on behalf of all authors\\
Technical University of Munich \& Helmholtz Centre Munich\\
\href{mailto:jinlong.ru@tum.de}{jinlong.ru@tum.de}

\bigskip
\noindent The Editors\\
\emph{Cell Press Multi-Journal Submission}

\bigskip
\noindent Dear Editors,

We are submitting our manuscript, \textbf{``A novel lytic \emph{Aeromonas}
bacteriophage rescues largemouth bass from lethal \emph{Aeromonas hydrophila}
infection,''} through \emph{Cell Press Multi-Journal Submission}, and we ask
that it be considered by \emph{Cell Reports} and \emph{iScience}.

\emph{Aeromonas hydrophila}, a leading pathogen of global freshwater
aquaculture, is increasingly multidrug-resistant, eroding antibiotic control
and intensifying the search for sustainable alternatives such as phage therapy.
Our study connects three findings:

\begin{enumerate}
\item \textbf{A highly distinct new viral genus.} Integrating nucleotide-level
(VIRIDIC, 83.65\% intergenomic similarity), whole-proteome (ViPTree),
amino-acid genome-BLAST distance phylogeny (VICTOR), and protein-sharing
network (vConTACT3) analyses, Hp3 and its closest relative HJ05 form a new,
highly distinct genus within the class \emph{Caudoviricetes}---with no more
than $S_G = 0.073$ whole-proteome similarity to any classified virus and clear
separation from all established families. This places Hp3 in an under-explored
region of viral sequence space and indicates that the phages of a globally
important aquaculture pathogen remain substantially undersampled.

\item \textbf{Demonstrated \emph{in vivo} efficacy and host safety.} A single
administration (MOI 1.0) increased the seven-day survival of largemouth bass
from 15\% to 85\% under lethal challenge, while fish receiving phage alone
showed 100\% survival with no histopathology.

\item \textbf{A genomically safe, well-defined biocontrol candidate.} Hp3
carries no antibiotic-resistance genes and no known virulence
determinants, and exhibits rapid lytic kinetics with broad temperature and pH
stability.

\end{enumerate}

We believe the work suits both journals. Its combination of a highly distinct
new viral lineage and a clear \emph{in vivo} therapeutic outcome offers the
broad, cross-disciplinary interest sought by \emph{Cell Reports}; and its
rigorous, multi-method genomic and \emph{in vivo} characterization---linking
the discovery of a genomically distinct virus to a measurable survival
benefit---fits the technically sound, multidisciplinary scope of
\emph{iScience}. We are happy to be routed to whichever journal the editors
judge the best fit.

The manuscript is original, has not been published previously, and is not under
consideration elsewhere. As corresponding author, I consent to its publication
and confirm that the content and authorship of this manuscript have been
approved by all authors; the authors declare no competing interests. All raw
sequencing data and assembled genomes are openly available in NCBI BioProject
\textbf{PRJNA1348731} (Hp3 genome under GenBank \textbf{PX754746.1}); figure
source data are deposited in Figshare
(\href{https://doi.org/10.6084/m9.figshare.30500525}{10.6084/m9.figshare.30500525})
and the analysis code is archived in Zenodo
(\href{https://doi.org/10.5281/zenodo.17454899}{10.5281/zenodo.17454899}).
All animal experiments followed the ARRIVE guidelines and were approved by the
Institutional Animal Care and Use Committee (approval no. NWAFU-314020038).

% --------------------------------------------------------------------------
% Suggested reviewers — five experts covering the three pillars of the paper.
% Substitute bench: Dann Turner (UWE Bristol, Dann2.Turner@uwe.ac.uk).
% --------------------------------------------------------------------------
\medskip
\noindent\textbf{Suggested reviewers} (none has a conflict of interest; none is a
recent collaborator or co-author of the authors):
\begin{itemize}
\item \textbf{Evelien M. Adriaenssens}, Quadram Institute Bioscience, Norwich, UK
(\href{mailto:evelien.adriaenssens@quadram.ac.uk}{evelien.adriaenssens@quadram.ac.uk})
--- bacteriophage genomics and prokaryotic virus taxonomy; Chair of the ICTV
Bacterial Viruses Subcommittee (\emph{Caudoviricetes} classification framework).
\item \textbf{Andrew D. Millard}, University of Leicester, UK
(\href{mailto:adm39@leicester.ac.uk}{adm39@leicester.ac.uk}) --- bacteriophage
comparative genomics, reference-genome curation, and therapeutic phages.
\item \textbf{Mao Lin}, Fisheries College, Jimei University, Xiamen, China
(\href{mailto:linmao@jmu.edu.cn}{linmao@jmu.edu.cn}) --- \emph{Aeromonas} and
aquaculture bacteriophage genomics and phage therapy (host-range and genome
characterization).
\item \textbf{Jianguo Su}, College of Fisheries, Huazhong Agricultural University,
Wuhan, China (\href{mailto:sujianguo@mail.hzau.edu.cn}{sujianguo@mail.hzau.edu.cn})
--- fish antibacterial immunity and \emph{A.~hydrophila} \emph{in vivo} challenge
models.
\item \textbf{Lingbing Zeng}, Yangtze River Fisheries Research Institute, Chinese
Academy of Fishery Sciences, Wuhan, China
(\href{mailto:zenglingbing@gmail.com}{zenglingbing@gmail.com}) --- fish pathology
and largemouth bass (\emph{Micropterus salmoides}) disease models.
\end{itemize}

\medskip
\noindent We thank you for considering our work and look forward to your
response.

\bigskip
\noindent Sincerely,\\[1ex]
Jinlong Ru, Li Deng, and Gaoxue Wang\\
\emph{(corresponding authors)}\\
on behalf of all authors

\end{document}
