---
title: "111-bacterial_genome"
date: today
params:
  name: "111-bacterial_genome"
---

**Updated: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'CET')` CET.**

```{r}
#| label: params
#| eval: !expr interactive()
#| include: false
params <- list(name = "111-bacterial_genome")
```

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false
#| cache: false
wd <- "analyses"
if (basename(getwd()) != wd) {
    setwd(here::here(wd))
}
here::i_am(paste0(params$name, ".qmd"), uuid = "247a7409-6777-4665-9620-fa40bd70ac76")
qproj::proj_create_dir_target(params$name, clean = FALSE)

# convenience helpers
path_target <- qproj::proj_path_target(params$name)
path_source <- qproj::proj_path_source(params$name)
path_raw <- path_source("00-raw")
path_resource <- here::here(path_raw, "d00-resource")
path_data <- here::here(path_raw, paste0("d", params$name))
dir.create(path_raw, recursive = TRUE, showWarnings = FALSE)
dir.create(path_data, recursive = TRUE, showWarnings = FALSE)
dir.create(path_resource, recursive = TRUE, showWarnings = FALSE)
```

## Tasks

1. The genome of *Aeromonas hydrophila* MS01 was assembled using [Unicycler](https://github.com/rrwick/Unicycler). 

2. Fragmented scaffolds were connected using [RagTag](https://github.com/malonge/RagTag) with *Aeromonas hydrophila* subsp. hydrophila ATCC 7966 as reference.

3. Assembled genome was annotated using [Bakta](https://github.com/oschwengers/bakta).

## Files written

These files have been written to the target directory, `r paste0("data/", params$name)`:

```{r}
#| label: list-files-target
knitr::kable(qproj::proj_dir_info(path_target(), tz = "CET"))
```
