---
title: "21-figs"
title-block-banner: true
author:
  - name: Jinlong Ru
    orcid: 0000-0002-6757-6018
date: 2025-10-23
toc: true
toc-depth: 4
number-sections: true
code-fold: true
code-line-numbers: true
code-tools: true
format: 
  html:
    embed-resources: true
    smooth-scroll: true
    page-layout: full
reference-location: section
citation-location: document
params:
  name: "21-figs"
---

**Updated: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'CET')` CET.**

The purpose of this document is to create figures based on the bacteriophage infection experiments.

```{r}
#| label: setup
#| message: false
#| include: false
#| warning: false
wd <- "analyses"
if (basename(getwd()) != wd) {
  setwd(here::here(wd))
}
params <- list(name = "21-figs")
here::i_am(paste0(params$name, ".qmd"), uuid = "a439e682-f799-4817-8772-5adc8157c4dd")
projthis::proj_create_dir_target(params$name, clean = FALSE)
path_target <- projthis::proj_path_target(params$name)
path_source <- projthis::proj_path_source(params$name)
path_raw <- path_source("00-raw")
path_resource <- here::here(path_raw, "d00-resource")
path_data <- here::here(path_raw, paste0("d", params$name))
dir.create(path_raw, recursive = TRUE, showWarnings = FALSE)
dir.create(path_data, recursive = TRUE, showWarnings = FALSE)
dir.create(path_resource, recursive = TRUE, showWarnings = FALSE)
```

```{r}
#| label: packages
#| message: false
suppressPackageStartupMessages({
  library(here)
  library(conflicted)
  library(tidyverse)
  library(data.table)
  library(openxlsx)
  library(svglite)
  devtools::load_all()
})

update_figure_theme <- function(plot) {
  plot + theme_classic(base_size = 13)
}

fpath_supp <- here("data/submit", "SupplementalData.xlsx")
```

## Tasks

### One-step growth curve

```{r}
df_onestep <- openxlsx::read.xlsx(fpath_supp, sheet = "growth_curve") %>% 
  dplyr::rowwise() %>%
  dplyr::mutate(mean_log = log10(mean(c(sample_1, sample_2, sample_3))),
         sd_log = sd(log10(c(sample_1, sample_2, sample_3)))
         )

plt_onestep <- ggplot(df_onestep, aes(x = time, y = mean_log)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_log - sd_log, ymax = mean_log + sd_log), width = 5, color = "black", size = 1) +
  scale_x_continuous(breaks = seq(0, 160, by = 15)) +
  labs(
    x = "Time (minutes)",
    y = expression(log[10]*"(PFU/mL)")
  )
```

### Temperature stability

```{r}
df_temperature <- openxlsx::read.xlsx(fpath_supp, sheet = "temperature_stability") %>% 
  dplyr::rowwise() %>%
  dplyr::mutate(mean_log = log10(mean(c(sample_1, sample_2, sample_3))),
         sd_log = sd(log10(c(sample_1, sample_2, sample_3)))
         )

plt_temperature <- ggplot(df_temperature, aes(x = temperature, y = mean_log)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_log - sd_log, ymax = mean_log + sd_log), width = 3, color = "black", size = 1) +
  scale_x_continuous(breaks = seq(0, 80, by = 10)) +
  scale_y_continuous(limits = c(1, 10)) +
  labs(
    x = "Temperature (°C)",
    y = expression(log[10]*"(PFU/mL)")
  )
```

### pH stability

```{r}
df_ph <- openxlsx::read.xlsx(fpath_supp, sheet = "pH_stability") %>% 
  dplyr::rowwise() %>%
  dplyr::mutate(mean_log = log10(mean(c(sample_1, sample_2, sample_3))),
         sd_log = sd(log10(c(sample_1, sample_2, sample_3)))
         )

plt_ph <- ggplot(df_ph, aes(x = pH, y = mean_log)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_log - sd_log, ymax = mean_log + sd_log), width = 0.2, color = "black", size = 1) +
  scale_x_continuous(breaks = seq(5, 12, by = 1), limits = c(5,11)) +
  scale_y_continuous(limits = c(1, 10)) + 
  labs(
    x = "pH",
    y = expression(log[10]*"(PFU/mL)")
  )
```

### MOI

```{r}
df_moi <- openxlsx::read.xlsx(fpath_supp, sheet = "MOI") %>% 
  rowwise() %>%
  mutate(
    mean_log = mean(log10(c(sample_1, sample_2, sample_3))),
    sd_log = sd(log10(c(sample_1, sample_2, sample_3)))
  ) %>% 
  ungroup() %>% 
  dplyr::mutate(MOI = factor(MOI, levels = c("0.001","0.01","0.1","1","10","100","1000")))


plt_moi <- ggplot(df_moi, aes(x = MOI, y = mean_log)) +
  geom_bar(stat = "identity", fill = "grey80", color = "black", width=0.7) +
  geom_errorbar(aes(ymin = mean_log - sd_log, ymax = mean_log + sd_log), width = 0.2, size=1) +
  labs(
    x = "MOI",
    y = expression(log[10]*"(PFU/mL)")
  )
```

### Combine figures together

Combine previous four figures into one single figure (2x2 grid)

```{r}
library(ggpubr)

fig_combined <- ggarrange(
  update_figure_theme(plt_moi),
  update_figure_theme(plt_onestep),
  update_figure_theme(plt_temperature),
  update_figure_theme(plt_ph),
  ncol = 2, nrow = 2,
  labels = c("A", "B", "C", "D"),
  font.label = list(size = 18, face = "bold")
)

# ggsave(path_target("combined_figures.svg"), plot = fig_combined, device = "svg", width = 10, height = 6.5)
ggsave(path_target("fig-combined.png"), plot = fig_combined, device = "png", width = 10, height = 6.5, dpi = 300)
```


### Lysis curve

```{r}
df_lysis <- openxlsx::read.xlsx(fpath_supp, sheet = "lysis_curve") %>% 
  reshape2::melt(id.vars = "time", variable.name = "sample_id", value.name = "OD600") %>% 
  dplyr::mutate(treatment = str_replace_all(sample_id, "_.*", ""))

df_lysis_summary <- df_lysis %>%
  group_by(treatment, time) %>%
  summarise(
    OD600_mean = mean(OD600, na.rm = TRUE),
    OD600_sd = sd(OD600, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    group_color = case_when(
      treatment == "ck" ~ "#464646",
      treatment == "moi0.01" ~ "#e94d61",
      treatment == "moi0.1" ~ "#399cf6",
      treatment == "moi1" ~ "#79e67e",
      treatment == "moi10" ~ "#9d5cd6",
      treatment == "moi100" ~ "#ebb85a",
      TRUE ~ "#000000"
    ),
    group_label = case_when(
      treatment == "ck" ~ "CK",
      treatment == "moi0.01" ~ "MOI 0.01",
      treatment == "moi0.1" ~ "MOI 0.1",
      treatment == "moi1" ~ "MOI 1",
      treatment == "moi10" ~ "MOI 10",
      treatment == "moi100" ~ "MOI 100",
      TRUE ~ "Other"
    )
  )

plt_lysis_summary <- ggplot(df_lysis_summary, aes(x = time, y = OD600_mean, color = group_label)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(ymin = OD600_mean - OD600_sd, ymax = OD600_mean + OD600_sd),
    width = 0.8, alpha = 0.9, linewidth = 0.7
  ) +
  scale_color_manual(
    values = unique(setNames(df_lysis_summary$group_color, df_lysis_summary$group_label))
  ) +
  labs(
    x = "Time (hours)",
    y = expression(OD[600]),
    color = NULL
  ) +
  # scale_x_continuous(breaks = seq(0, 48, 5), expand = c(0,0)) +
  # scale_y_continuous(limits = c(0, 0.8), expand = c(0,0)) +
  annotate("text", x = 33, y = 0.15, label = "ANOVA (48h): p < 0.001", 
           hjust = 0, size = 4, fontface = "italic") +
  theme_classic(base_size = 13) +
  theme(
    legend.position = c(0.11, 0.78),
    legend.background = element_rect(fill = "white", color = "gray80")
  )

plt_lysis_summary

ggsave(path_target("fig3.png"), plot = plt_lysis_summary, device = "png", width = 6.5, height = 4.5, dpi = 300)
```

```{r}
# --- Add this to your R code to calculate Lysis Stats ---

# 1. Select the final time point (e.g., 48h) or a specific time point of interest
lysis_endpoint <- df_lysis %>% 
  dplyr::filter(time == 48) # Assuming 48 is the last time point

# 2. Run One-way ANOVA
anova_res <- aov(OD600 ~ treatment, data = lysis_endpoint)
summary_anova <- summary(anova_res)

# Extract values
f_value <- summary_anova[[1]][["F value"]][1]
df_num <- summary_anova[[1]][["Df"]][1]
df_den <- summary_anova[[1]][["Df"]][2]
p_val_anova <- summary_anova[[1]][["Pr(>F)"]][1]

cat(sprintf("Lysis ANOVA (48h): F(%d, %d) = %.2f, p-value = %.3f\n", 
            df_num, df_den, f_value, p_val_anova))

cat(sprintf("Lysis ANOVA (48h): F(%d, %d) = %f, p-value = %f\n", 
            df_num, df_den, f_value, p_val_anova))

df_num
df_den
f_value
p_val_anova
```



### Survival curve

```{r}
df_survival <- openxlsx::read.xlsx(fpath_supp, sheet = "survival_curve") %>% 
  reshape2::melt(id.vars = "time", variable.name = "Group", value.name = "Count") %>% 
  mutate(
    survival_rate = Count / 20 * 100,
    group_color = case_when(
      Group == "PBS_group" ~ "#2166ac",
      Group == "phage_group" ~ "#1a9850",
      Group == "bacterial_group" ~ "#ef8a62",
      Group == "treatment_group" ~ "#f637e0"
    ),
    group_shape = case_when(
      Group == "PBS_group" ~ 18,
      Group == "phage_group" ~ 15,
      Group == "bacterial_group" ~ 17,
      Group == "treatment_group" ~ 16
    ),
    group_linetype = case_when(
      Group == "PBS_group" ~ "dotted",
      Group == "phage_group" ~ "solid",
      Group == "bacterial_group" ~ "dotted",
      Group == "treatment_group" ~ "solid"
    ),
    group_label = case_when(
      Group == "PBS_group" ~ "PBS group",
      Group == "phage_group" ~ "Phage group",
      Group == "bacterial_group" ~ "Bacterial group",
      Group == "treatment_group" ~ "Treatment group"
    )
  )

plt_survival <- ggplot(df_survival, aes(x = time, y = survival_rate, group = Group)) +
  geom_line(aes(color = group_label), size = 1, alpha = 0.4) +
  geom_point(aes(color = group_label, shape = group_label), size = 5, stroke = 1, alpha = 0.6) +
  scale_color_manual(values = setNames(df_survival$group_color, df_survival$group_label)) +
  scale_shape_manual(values = setNames(df_survival$group_shape, df_survival$group_label)) +
  scale_x_continuous(breaks = seq(min(df_survival$time), max(df_survival$time), by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 105), expand = c(0, 0)) +
  labs(
    x = "Time (days)",
    y = "Survival rate (%)"
  ) +
  annotate("text", x = 0, y = 10, label = "Log-rank test, p < 0.001", 
           hjust = 0, size = 4, fontface = "italic") +
  theme_classic(base_size = 13) +
  theme(
    legend.position = c(0.7, 0.45),
    legend.background = element_rect(fill = "white", color = "gray60"),
    legend.title = element_blank()
  )

plt_survival

ggsave(path_target("fig-survival.png"), plt_survival, device = "png", width = 6, height = 4, dpi = 300)
```

```{r}
# --- Add this to your R code to calculate Survival Stats ---

library(survival)

# 1. Reconstruct individual-level data from the aggregate counts for survival analysis
# We need a dataframe with: Time, Status (1=dead, 0=censored), Group
reconstruct_survival_data <- function(df_counts) {
  # Total fish per group = 20
  total_n <- 20
  
  # Filter only the groups we want to compare (e.g., Treatment vs Bacterial)
  target_groups <- c("bacterial_group", "treatment_group")
  df_sub <- df_counts %>% dplyr::filter(Group %in% target_groups)
  
  all_data <- data.frame()
  
  for (grp in unique(df_sub$Group)) {
    grp_data <- df_sub %>% dplyr::filter(Group == grp) %>% dplyr::arrange(time)
    
    # Calculate deaths at each time point
    # Initial count is 20. If day 1 count is 18, then 2 died at day 1.
    current_survivors <- total_n
    deaths_vector <- c()
    times_vector <- c()
    
    for (i in 1:nrow(grp_data)) {
      day <- grp_data$time[i]
      survivors <- grp_data$Count[i]
      n_deaths <- current_survivors - survivors
      
      if (n_deaths > 0) {
        deaths_vector <- c(deaths_vector, rep(day, n_deaths))
      }
      current_survivors <- survivors
    }
    
    # The rest are censored at the end of experiment (Day 7)
    censored_count <- current_survivors
    censored_vector <- rep(7, censored_count) # Assuming exp ends at day 7
    
    # Combine
    time_points <- c(deaths_vector, censored_vector)
    status <- c(rep(1, length(deaths_vector)), rep(0, length(censored_vector))) # 1=dead
    
    temp_df <- data.frame(
      time = time_points,
      status = status,
      group = grp
    )
    all_data <- rbind(all_data, temp_df)
  }
  return(all_data)
}

surv_data_for_stats <- reconstruct_survival_data(df_survival)

# 2. Run Log-rank test (Mantel-Cox)
surv_diff <- survdiff(Surv(time, status) ~ group, data = surv_data_for_stats)

print(surv_diff)
# Extract p-value and Chi-sq manually to print cleanly
p_val_surv <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n) - 1)
cat(sprintf("Survival Log-rank Test: Chi-sq = %f, df = %d, p-value = %f\n", 
            surv_diff$chisq, length(surv_diff$n) - 1, p_val_surv))
```

## Files written

These files have been written to the target directory, `r paste0("data/", params$name)`:

```{r}
#| label: list-files-target
projthis::proj_dir_info(path_target(), tz = "CET")
```
