---
title: "404-safety"
date: today
params:
  name: "404-safety"
---

**Updated: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'CET')` CET.**

```{r}
#| label: params
#| eval: !expr interactive()
#| include: false
params <- list(name = "404-safety")
```

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false
#| cache: false
wd <- "analyses"
if (basename(getwd()) != wd) {
    setwd(here::here(wd))
}
here::i_am(paste0(params$name, ".qmd"), uuid = "f9a04ec2-3586-4c25-9fcd-2f22f7c151ce")
qproj::proj_create_dir_target(params$name, clean = FALSE)

# convenience helpers
path_target <- qproj::proj_path_target(params$name)
path_source <- qproj::proj_path_source(params$name)
path_raw <- path_source("00-raw")
path_resource <- here::here(path_raw, "d00-resource")
path_data <- here::here(path_raw, paste0("d", params$name))
dir.create(path_raw, recursive = TRUE, showWarnings = FALSE)
dir.create(path_data, recursive = TRUE, showWarnings = FALSE)
dir.create(path_resource, recursive = TRUE, showWarnings = FALSE)
```

```{r}
#| label: packages
#| message: false
#| warning: false
#| cache: false
suppressPackageStartupMessages({
  library(here)
  library(conflicted)
  library(tidyverse)
  library(data.table)
  devtools::load_all()
})
# Resolve common dplyr / data.table / stats name clashes up front.
# Add more `conflicts_prefer()` calls below as needed.
conflicted::conflicts_prefer(
  dplyr::filter,
  dplyr::lag,
  dplyr::first,
  dplyr::last,
  dplyr::between,
  .quiet = TRUE
)
```

## Tasks

> **共享底座 step 4/6 · 治疗安全筛查** ⭐线②核心卖点。映射 `shared-base-spec §4 ⭐②` + `line2-db-schema §1 safety 表`。
> 注释结果 → 毒力/ARG/溶原模块判定 → safety_grade，交 `406-master_table` + 线②。

**上游**：`path_source("403-annotate", ...)`。**外部工具**：VFDB(毒力) · CARD/AMRFinderPlus/ResFinder(ARG) · 溶原模块(integrase/recombinase/repressor) · toxin。

**判定规则**(line2-db-schema §1)：有 ARG 或 毒力 或 溶原模块(integrase+repressor) → caution/unsafe；全清 → safe(治疗候选)。

### Task 1: 毒力 / ARG / 溶原模块扫描

```{r}
#| label: safety-scan
#| eval: false
# TODO[404-T1]: VFDB + CARD/AMRFinderPlus + integrase/repressor/toxin HMM
```

### Task 2: safety_grade 计算

```{r}
#| label: safety-grade
#| eval: false
# TODO[404-T2]: 按规则定 safe/caution/unsafe
# write: path_target("safety.tsv")
```

## Files written

These files have been written to the target directory, `r paste0("data/", params$name)`:

```{r}
#| label: list-files-target
knitr::kable(qproj::proj_dir_info(path_target(), tz = "CET"))
```
